### 2019-05-23 README.Rmd
discussion

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below. 
Title <- "Territoriality and the organization of technology during the Last Glacial Maximum in southwestern Europe"

```

[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-brightgreen.svg)](https://github.com/jmcascalheira/ScaledPiecesVB/commits/master)
[![minimal R version](https://img.shields.io/badge/R%3E%3D-3.2.4-brightgreen.svg)](https://cran.r-project.org/) 
[![Build Status](https://travis-ci.org/jmcascalheira/ScaledPiecesVB.svg?branch=master)](https://travis-ci.org/jmcascalheira/ScaledPiecesVB)
[![Licence](https://img.shields.io/github/license/mashape/apistatus.svg)](http://choosealicense.com/licenses/mit/)
[![ORCiD](https://img.shields.io/badge/ORCiD-/0000--0003--0321--8892-green.svg)](http://orcid.org/0000-0003-0321-8892)

# LGMTechnology

This repository contains the data and code for our paper:

> Cascalheira, J. (2019). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx>

Our pre-print is online here:

> Cascalheira, J. (2019). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to cite

Please cite this compendium as:

> Cascalheira, J. (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to download or install

You can download the compendium as a zip from from this URL: </archive/master.zip>

Or you can install this compendium as an R package, LGMTechnology, from GitHub with:

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

I welcome contributions from everyone. Before you get started, please see my [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.