### 2019-05-23 analysis/paper/elongated_MCA.Rmd

---
title: "Elongated products"
author: 
  - name: João Cascalheira
    email: jmcascalheira@ualg.pt
    footnote: Corresponding author

address: 
  - code: ICArEHB
    address: | 
      FCHS, 
      Universidade do Algarve,
      Campus de Gambelas,
      8005-139 Faro, Portugal 
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: 
    rmarkdown::html_document:
        number_sections: yes
        toc: yes

---

```{r setup, include = FALSE, echo = FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>"
  )

library(knitr)
library(rmarkdown)
library(git2r)
library(readr)
library(dplyr)
library(tidyr)
library(tibble)
library(tab)
library(forcats)
library(FactoMineR)
library(factoextra)
library(multcompView)
library(RcmdrMisc)
library(WRS2)

```

```{r load_data, echo = FALSE}

# Load data, calculate Elongation/Flattening for blanks and tidy data (including lumping categories with little (5%) representation into 'Other')

blanks <- read.csv("../data/raw_data/blanks.csv")

elongated_tidy <- blanks %>%
  filter(Class == "Elongated product") %>%
  drop_na(Length, Width, Thickness) %>%
  mutate(Elongation = Length / Width) %>%
  mutate(Flattening = Width / Thickness) %>%
  mutate(Profile = fct_lump(Profile, prop = 0.05)) %>%
  mutate(EdgeShape = fct_lump(EdgeShape, prop = 0.05)) %>%
  mutate(CrossSection = fct_lump(CrossSection, prop = 0.05)) %>%
  mutate(PlatformType = fct_lump(PlatformType, prop = 0.05)) %>%
  mutate(CortexLoc = fct_lump(CortexLoc, prop = 0.05)) %>%
  mutate(DorsalPattern = fct_lump(DorsalPattern, prop = 0.05)) %>%
  mutate(Termination = fct_lump(Termination, prop = 0.05)) %>%
  mutate(CortexPerc = dplyr::recode(CortexPerc, "75-95%" = "76-100%", "25-75%" ="26-75%", "100%" = "76-100%", "<25%" = "1-25%", ">95%" = "76-100%", "0%" = "0%")) %>%
  mutate(ScarNumber = as.factor(ScarNumber)) %>%
  mutate(DorsalPattern = dplyr::recode(DorsalPattern, "Unidentifiable" = "Other")) %>%
  mutate(ScarNumber = dplyr::recode(ScarNumber, "4" = "4 or more", "5" = "4 or more", "6" = "4 or more", "7" = "4 or more", "8" = "4 or more", "9" = "4 or more"))

# Rename Context column
colnames(elongated_tidy)[1] <- "Context"

# Convert ScarNumber to factor
elongated_tidy$ScarNumber <- factor(elongated_tidy$ScarNumber)

# Bin continous variables ("natural" is used to cut between bins to be determined by a k-means clustering)
elongated_tidy$ElongFact <- binVariable(elongated_tidy$Elongation, bins = 2, method = "natural", labels = c("low", "high"))
elongated_tidy$FlattFact <- binVariable(elongated_tidy$Flattening, bins = 2, method = "natural", labels = c("low", "high"))

```

```{r elongated_indices, echo=FALSE}

ggplot(elongated_tidy, aes(x=Width, y=Length, color=ElongFact)) + geom_point()

ggplot(elongated_tidy, aes(x=Thickness, y=Width, color=FlattFact)) + geom_point()

```

```{r reduction_plots, echo=FALSE}

reduction <- blanks %>%
  drop_na(Length, Width, Thickness) %>%
  mutate(Area = Width * Length) %>%
  mutate(CortexPerc = dplyr::recode(CortexPerc, "75-95%" = "76-100%", "25-75%" ="26-75%", "100%" = "76-100%", "<25%" = "1-25%", ">95%" = "76-100%", "0%" = "0%"))

colnames(reduction)[1] <- "Context"

reduction <- reduction %>%   
  dplyr::select(Context, CortexPerc, Area) 
  #mutate(Context = dplyr::recode(Context, "AMB II" = "AMB", "AMB IV" = "AMB", "AMB VI" = "AMB", "PAP 4'00-4'75" = "PAP",
  #                               "PAP 4'75-5'25" = "PAP", "PAP 5'25-6'25" = "PAP", "VB A" = "VB",
  #                              "VB B" = "VB", "VB C" = "VB"))

reduction$ScarNumber <- as.numeric(reduction$ScarNumber)

reduction$CortexPerc <- factor(reduction$CortexPerc, levels = c("0%","1-25%","26-75%","76-100%"))

data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}

reduction_plot <- data_summary(reduction, varname="Area", 
                    groupnames=c("Context", "CortexPerc"))

ggplot(reduction_plot, aes(x=CortexPerc, y=Area, group=Context, color=Context)) + 
  geom_line() +
  geom_point() +
  theme_classic() +
  labs(x = "Cortex %")

```

```{r elongated_histogram, echo = FALSE}

### Plot width histograms for elongated products
blades <- blanks_tidy %>%
  filter(Class == "Elongated product") +
  ggplot(blades, aes(Width)) +
  geom_histogram(aes(y = ..density..)) +
  geom_density() +
  facet_wrap(~Context, ncol = 2) +
  scale_y_continuous(labels = scales::percent)

```

```{r elongated_cross_table, echo=FALSE}

elongated_table <- elongated_tidy %>% 
  filter(Class == "Elongated product") %>% 
  droplevels()

# Set variables
var_list <- c("PlatformType", "CrossSection", "Profile", "DorsalPattern", "EdgeShape", "Elongation", "Flattening")

# Get cross-tables by chronology using cross_tb() function
elongated_table <- tabmulti(elongated_table, "Context", var_list,
                 p.include = FALSE,
                 n.headings = FALSE,
                 means.text.label = "none",
                 freq.text.label = "none")

elongated_table <- elongated_table %>% 
  as_tibble() %>% 
  select(" " = Variable, `AMB II`, `AMB IV`, `AMB VI`, `PAP 4'00-4'75`, `PAP 4'75-5'25`, `PAP 5'25-6'25`, `VALM`, `VB A`, `VB B`, `VB C`, Total = Overall)

kable(elongated_table)

```

```{r MCA_core_exploitation_elongated, echo=FALSE}

# Filter specific categories for MCA analysis
core_exploitation_elongated <- elongated_tidy %>%
  filter(CortexPerc != "76-100%") %>%
  filter(ScarNumber != "0") %>%
  dplyr::select(Context, CortexPerc, DorsalPattern, ElongFact, ScarNumber) %>%
  na.omit() %>%
  droplevels()

# Run MCA

mca_res_core_exploitation_elongated <- MCA(core_exploitation_elongated, graph = FALSE, quali.sup = 1, method = "Burt")

fviz_screeplot(mca_res_core_exploitation_elongated, addlabels = TRUE, ylim = c(0, 45))

fviz_mca_var(mca_res_core_exploitation_elongated, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE, invisible = "quali.sup",
             ggtheme = theme_gray())

```

# Variable contribution to MCA solution

```{r}

var_description_core_exploitation_elongated <- data.frame(
  Dimension = c("Dimension 1", "Dimension 2"),
  Variability = c("28.9%", "16.5%"),
  Positive = c("Production of elongated blanks from cores with high amounts of cortex and little amount of removals on debitage surfaces, using only unidireccional strategies",
    "Use of bidireccional strategies to produce elongated products from cores with some cortical surface on debitage plane, but with complex removal biography"),
  Negative = c("Removal of highly elongated blanks using bidireccional strategies from cores with no cortex on debitage surface and complex, highly developed, reduction history",
  "Use of unidireccional or other reduction strategies to produce elongated blanks with small amounts of cortex and few removals on dorsal surfaces"
))

kable(var_description_core_exploitation_elongated)

```

```{r ANOVA_comparisons, echo = FALSE}

# List of comparisons to use with grouping labels (dash with no spaces must be the separator)
comparisons <- c("AMB II-AMB IV", "AMB II-AMB VI", "AMB II-PAP 4'00",
                 "AMB II-PAP 4'75", "AMB II-PAP 5'25", "AMB II-VALM",
                 "AMB II-VB A", "AMB II-VB B", "AMB II-VB C",
                 "AMB IV-AMB VI", "AMB IV-PAP 4'00", "AMB IV-PAP 4'75",
                 "AMB IV-PAP 5'25", "AMB IV-VALM", "AMB IV-VB A", "AMB IV-VB B",
                 "AMB IV-VB C", "AMB VI-PAP 4'00", "AMB VI-PAP 4'75",
                 "AMB VI-PAP 5'25", "AMB VI-VALM", "AMB VI-VB A",
                 "AMB VI-VB B", "AMB VI-VB C", "PAP 4'00-PAP 4'75",
                 "PAP 4'00-PAP 5'25", "PAP 4'00-VALM", "PAP 4'00-VB A",
                 "PAP 4'00-VB B", "PAP 4'00-VB C", "PAP 4'75-PAP 5'25",
                 "PAP 4'75-VALM", "PAP 4'75-VB A", "PAP 4'75-VB B",
                 "PAP 4'75-VB C", "PAP 5'25-VALM", "PAP 5'25-VB A",
                 "PAP 5'25-VB B", "PAP 5'25-VB C", "VALM-VB A", "VALM-VB B",
                 "VALM-VB C", "VB A-VB B", "VB A-VB C", "VB B-VB C")

```

```{r ANOVA_core_exploitation_elongated, echo=FALSE}

## From script tukey_boxplot

# Function below does not allow hyphen between categories

mca <- as.data.frame(mca_res_core_exploitation_elongated$ind)
context <- as.data.frame(core_exploitation_elongated$Context)

mca <- mca %>%
  dplyr::select(1:2)

mca <- bind_cols(context, mca)

mca <- mca %>%
  mutate(context = `core_exploitation_elongated$Context`, "dim1" = coord.Dim.1, "dim2" = coord.Dim.2) %>%
  dplyr::select(context, dim1, dim2)

mca <- mca %>%
  mutate(context = dplyr::recode(context, "PAP 5'25-6'25" = "PAP 5'25 6'25",
                                 "PAP 4'75-5'25" = "PAP 4'75 5'25",
                                 "PAP 4'00-4'75" = "PAP 4'00 4'75"))

# Run bootstrap ANOVA
boot_anova <- t1waybt(dim1 ~ context, data = mca, tr=.2, nboot=5000)

# Multi-comparison bootstrap
multi_comp <- mcppb20(dim1 ~ context, data = mca, tr=.2, nboot=5000)

# Convert results to data frame
multi_comp <- as.data.frame(multi_comp$comp)

# Adjust p-values using FDR method
#p.value.adjusted <- p.adjust(multi_comp[,6], "fdr")

# Combine with data frame
#multi_comp <- cbind(multi_comp, p.value.adjusted)

# Add contexts names and transform to column "context"
row.names(multi_comp) <- comparisons
multi_comp <- rownames_to_column(multi_comp, "context")

# Boxplot with representation of grouping calculated by multicomparison analysis

# Build list with pairwise p-values
LABELS <- setNames(multi_comp$`p-value`, multi_comp$context)

# Attribute letters to each grouping based on alpha = .05
grouping <- data.frame(multcompLetters(LABELS)['Letters'])
grouping <- rownames_to_column(grouping, "context")

# Colors for boxplot
my_colors <- c( rgb(143,199,74,maxColorValue = 255),
                rgb(242,104,34,maxColorValue = 255),
                rgb(111,145,202,maxColorValue = 255),
                rgb(254,188,18,maxColorValue = 255),
                rgb(74,132,54,maxColorValue = 255),
                rgb(236,33,39,maxColorValue = 255),
                rgb(165,103,40,maxColorValue = 255))

# Start boxplot
DIM1_boxplot <- boxplot(mca$dim1 ~ mca$context , ylim=c(min(mca$dim1),
                        1.1*max(mca$dim1)), col=my_colors[as.numeric(grouping[,2])],
                        ylab="Dimension 1" , main="", outline = FALSE)
# Location of labels
over <- 0.1*max(DIM1_boxplot$stats[nrow(DIM1_boxplot$stats),] )

# Superimpose labels
text( c(1:nlevels(mca$contex)) , DIM1_boxplot$stats[nrow(DIM1_boxplot$stats),]+over , grouping[,2]  , col=my_colors[as.numeric(grouping[,2])] )

```

```{r MCA_platform_elongated, echo=FALSE}

# Filter specific categories for MCA analysis
platform_elongated <- elongated_tidy %>%
  filter(CortexPerc != "76-100%") %>%
  dplyr::select(Context, PlatformType, CortexPerc, ElongFact) %>%
  na.omit() %>%
  droplevels()

# Run MCA

mca_res_platform_elongated <- MCA(platform_elongated, graph = FALSE, quali.sup = 1, method = "Burt")

fviz_screeplot(mca_res_platform_elongated, addlabels = TRUE, ylim = c(0, 45))

fviz_mca_var(mca_res_platform_elongated, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE, invisible = "quali.sup",
             ggtheme = theme_gray(), title = "")

```

# Variable contribution to MCA solution

```{r}

var_description_platform_elongated <- data.frame(
  Dimension = c("Dimension 1", "Dimension 2"),
  Variability = c("18.3%", "14.1%"),
  Positive = c("Use of cortical platforms to extract products with low elongation with some cortex on dorsal surface",
    "Use of cortical platforms to extract highly elongated blanks from cores with small amounts of cortex on the debitage surface"),
  Negative = c("Use of plainplatforms to produce highly elongated blanks from completely decorticated debitage surfaces",
  "Production of blanks with low elongation, from cores with some cortex surface and faceted platforms"
))

kable(var_description_platform_elongated)

```

```{r ANOVA_platform_elongated, echo=FALSE}

# Prepare data for bootstraped ANOVA

mca <- as.data.frame(mca_res_platform_elongated$ind)
context <- as.data.frame(platform_elongated$Context)

mca <- mca %>%
  dplyr::select(1:2)

mca <- bind_cols(context, mca)

mca <- mca %>%
  mutate(context = `platform_elongated$Context`, "dim1" = coord.Dim.1, "dim2" = coord.Dim.2) %>%
  dplyr::select(context, dim1, dim2)

mca <- mca %>%
  mutate(context = dplyr::recode(context, "PAP 5'25-6'25" = "PAP 5'25 6'25",
                                 "PAP 4'75-5'25" = "PAP 4'75 5'25",
                                 "PAP 4'00-4'75" = "PAP 4'00 4'75"))

# Run bootstrap ANOVA
boot_anova <- t1waybt(dim1 ~ context, data = mca, tr=.2, nboot=5000)

# Multi-comparison bootstrap
multi_comp <- mcppb20(dim1 ~ context, data = mca, tr=.2, nboot=5000)

# Convert results to data frame
multi_comp <- as.data.frame(multi_comp$comp)

# Adjust p-values using FDR method
#p.value.adjusted <- p.adjust(multi_comp[,6], "fdr")

# Combine with data frame
#multi_comp <- cbind(multi_comp, p.value.adjusted)

# Add contexts names and transform to column "context"
row.names(multi_comp) <- comparisons
multi_comp <- rownames_to_column(multi_comp, "context")

# Boxplot with representation of grouping calculated by multicomparison analysis

# Build list with pairwise p-values
LABELS <- setNames(multi_comp$`p-value`, multi_comp$context)

# Attribute letters to each grouping based on alpha = .05
grouping <- data.frame(multcompLetters(LABELS)['Letters'])
grouping <- rownames_to_column(grouping, "context")

# Colors for boxplot
my_colors <- c( rgb(143,199,74,maxColorValue = 255),
                rgb(242,104,34,maxColorValue = 255),
                rgb(111,145,202,maxColorValue = 255),
                rgb(254,188,18,maxColorValue = 255),
                rgb(74,132,54,maxColorValue = 255),
                rgb(236,33,39,maxColorValue = 255),
                rgb(165,103,40,maxColorValue = 255))

# Start boxplot
DIM1_boxplot <- boxplot(mca$dim1 ~ mca$context , ylim=c(min(mca$dim1),
                        1.1*max(mca$dim1)), col=my_colors[as.numeric(grouping[,2])],
                        ylab="Dimension 1" , main="", outline = FALSE)
# Location of labels
over <- 0.1*max(DIM1_boxplot$stats[nrow(DIM1_boxplot$stats),] )

# Superimpose labels
text( c(1:nlevels(mca$contex)) , DIM1_boxplot$stats[nrow(DIM1_boxplot$stats),]+over , grouping[,2]  , col=my_colors[as.numeric(grouping[,2])] )

```