### 2019-09-18 analysis/paper/cores_MCA.Rmd

---
title: "Elongated products"
author: 
  - name: João Cascalheira
    email: jmcascalheira@ualg.pt
    footnote: Corresponding author

address: 
  - code: ICArEHB
    address: | 
      FCHS, 
      Universidade do Algarve,
      Campus de Gambelas,
      8005-139 Faro, Portugal 
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: 
    rmarkdown::html_document:
        number_sections: yes
        toc: yes

---

```{r setup, include = FALSE, echo = FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>"
  )

library(knitr)
library(rmarkdown)
library(git2r)
library(readr)
library(dplyr)
library(tidyr)
library(tibble)
library(tab)
library(forcats)
library(FactoMineR)
library(factoextra)
library(multcompView)
library(RcmdrMisc)

```

```{r load_data, echo = FALSE}

# Load data, calculate Elongation/Flattening for blanks and tidy data (including lumping categories with little (5%) representation into 'Other')

cores <- read.csv("../data/raw_data/cores.csv")

cores_tidy <- cores %>%
  drop_na(Length, Width, Thickness, Weight) %>%
  mutate(Elongation = Length / Width) %>%
  mutate(Flattening = Width / Thickness) %>%
  mutate(CoreSection = fct_lump(CoreSection, prop = 0.05)) %>%
  mutate(PlatformType = fct_lump(PlatformType, prop = 0.05)) %>%
  mutate(CortexLoc = fct_lump(CortexLoc, prop = 0.05)) %>%
  mutate(CoreType = fct_lump(CoreType, prop = 0.05)) %>%
  mutate(BlankType = fct_lump(BlankType, prop = 0.05)) %>%
  mutate(Platforms = fct_lump(Platforms, prop = 0.05)) %>%
  mutate(CortexPerc = dplyr::recode(CortexPerc, "75-95%" = "76-100%", "25-75%" ="26-75%", "100%" = "76-100%", "<25%" = "1-25%", ">95%" = "76-100%", "0%" = "0%")) %>%
  mutate(ScarNumber = as.factor(ScarNumber)) %>%
  mutate(ScarNumber = dplyr::recode(ScarNumber, "4" = "4 or more", "5" = "4 or more", "6" = "4 or more", "7" = "4 or more", "8" = "4 or more", "9" = "4 or more"))

# Rename Context column
colnames(cores_tidy)[1] <- "Context"

# Bin continous variables
cores_tidy$ElongFact <- binVariable(cores_tidy$Elongation, bins = 3, method = "natural", labels = c("low", "medium", "high"))
cores_tidy$FlattFact <- binVariable(cores_tidy$Flattening, bins = 3, method = "natural", labels = c("low", "medium", "high"))
cores_tidy$WeightFact <- binVariable(cores_tidy$Weight, bins = 3, method = "natural", labels = c("low", "medium", "high"))
```

```{r flake_cross_table, echo=FALSE}

cores_table <- cores_tidy %>%
  droplevels()

# Set variables
var_list <- c("CoreType", "Platforms", "CoreSection", "BlankType", "CortexPerc", "Weight",
              "Elongation", "Flattening")

# Get cross-tables by chronology using cross_tb() function
cores_table <- tabmulti(cores_table, "Context", var_list,
                 p.include = FALSE,
                 n.headings = FALSE,
                 means.text.label = "none",
                 freq.text.label = "none")

cores_table <- cores_table %>% 
  as_tibble() %>% 
  select(" " = Variable, `AMB II`, `AMB IV`, `AMB VI`, `PAP 4'00-4'75`, `PAP 4'75-5'25`, `PAP 5'25-6'25`, `VALM`, `VB A`, `VB B`, `VB C`, Total = Overall)

kable(cores_table)

```

```{r MCA_cores, echo=FALSE}

# Filter specific categories for MCA analysis
cores <- cores_tidy %>%
  select(Context, CortexPerc, CoreType, Platforms, ElongFact, FlattFact, BlankType, PlatformType) %>%
  na.omit() %>%
  droplevels()

# Run MCA

mca_res_cores <- MCA(cores, graph = FALSE, quali.sup = 1, method = "Burt")

fviz_screeplot(mca_res_cores, addlabels = TRUE, ylim = c(0, 45))

fviz_mca_var(mca_res_cores, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE, invisible = "quali.sup",
             ggtheme = theme_gray())

```

# Variable contribution to MCA solution

```{r}

var_description <- data.frame(
  Dimension = c("Dimension 1", "Dimension 2"),
  Variability = c("23%", "13%"),
  Positive = c("Simple cores with 2 isolated or multiple platforms. Low elongation and high flattening. Extraction of flakes with cortical platforms",
    "Other types of cores with multiple platforms for mixed products"),
  Negative = c("Prismatic cores with 2 opposed platforms. High elongation and low flattening, for elongated products",
  "Simple cores with 1 platform and high % of cortex. Extraction of blades with cortical platforms"
))

kable(var_description)

```

```{r ANOVA_cores, echo=FALSE}

## From script tukey_boxplot

# Function below does not allow hyphen between categories

mca <- as.data.frame(mca_res_cores$ind)
context <- as.data.frame(cores_tidy$Context)

mca <- mca %>%
  select(1:2)

mca <- bind_cols(context, mca)

mca <- mca %>%
  mutate(context = `cores_tidy$Context`, "dim1" = coord.Dim.1, "dim2" = coord.Dim.2) %>%
  select(context, dim1, dim2)

mca <- mca %>%
  mutate(context = dplyr::recode(context, "PAP 5'25-6'25" = "PAP 5'25 6'25",
                                 "PAP 4'75-5'25" = "PAP 4'75 5'25",
                                 "PAP 4'00-4'75" = "PAP 4'00 4'75"))

# DIMENSION 1

model <- lm(mca$dim1 ~ mca$context)
fit <- aov(model)
TUKEY <- TukeyHSD(x=fit, 'mca$context', conf.level=0.95)

generate_label_df <- function(TUKEY, variable){

# Extract labels and factor levels from Tukey post-hoc
Tukey.levels <- TUKEY[[variable]][,4]
Tukey.labels <- data.frame(multcompLetters(Tukey.levels)['Letters'])

# Put the labels in the same order as in the boxplot :
Tukey.labels$context=rownames(Tukey.labels)
Tukey.labels=Tukey.labels[order(Tukey.labels$context) , ]
return(Tukey.labels)
}

LABELS <- generate_label_df(TUKEY , "mca$context")

my_colors <- c( rgb(143,199,74,maxColorValue = 255),
             rgb(242,104,34,maxColorValue = 255),
             rgb(111,145,202,maxColorValue = 255),
             rgb(254,188,18,maxColorValue = 255),
             rgb(74,132,54,maxColorValue = 255),
             rgb(236,33,39,maxColorValue = 255),
             rgb(165,103,40,maxColorValue = 255))

# Draw the basic boxplot
DIM1_boxplot <- boxplot(mca$dim1 ~ mca$context , ylim=c(min(mca$dim1),
                                          1.1*max(mca$dim1)),
          col=my_colors[as.numeric(LABELS[,1])] , ylab="Dimesion 1" , main="", outline = FALSE)

# I want to write the letter over each box. Over is how high I want to write it.
over <- 0.1*max(DIM1_boxplot$stats[nrow(DIM1_boxplot$stats),] )

#Add the labels
text( c(1:nlevels(mca$contex)) , DIM1_boxplot$stats[nrow(DIM1_boxplot$stats),]+over , LABELS[,1]  , col=my_colors[as.numeric(LABELS[,1])] )

# DIMENSION 2

model <- lm(mca$dim2 ~ mca$context)
fit <- aov(model)
TUKEY <- TukeyHSD(x=fit, 'mca$context', conf.level=0.95)

generate_label_df <- function(TUKEY, variable){

# Extract labels and factor levels from Tukey post-hoc
Tukey.levels <- TUKEY[[variable]][,4]
Tukey.labels <- data.frame(multcompLetters(Tukey.levels)['Letters'])

# Put the labels in the same order as in the boxplot
  Tukey.labels$context=rownames(Tukey.labels)
  Tukey.labels=Tukey.labels[order(Tukey.labels$context) , ]
  return(Tukey.labels)
}

LABELS <- generate_label_df(TUKEY , "mca$context")

my_colors <- c(rgb(143,199,74,maxColorValue = 255),
             rgb(242,104,34,maxColorValue = 255),
             rgb(111,145,202,maxColorValue = 255),
             rgb(254,188,18,maxColorValue = 255),
             rgb(74,132,54,maxColorValue = 255),
             rgb(236,33,39,maxColorValue = 255),
             rgb(165,103,40,maxColorValue = 255))

# Draw the basic boxplot
DIM2_boxplot <- boxplot(mca$dim2 ~ mca$context , ylim=c(min(mca$dim2),
                                          1.1*max(mca$dim2)),
          col=my_colors[as.numeric(LABELS[,1])] , ylab="Dimension 2" , main="", outline = FALSE)

# Write the letter over each box. Over is how high I want to write it.
over <- 0.1*max(DIM2_boxplot$stats[nrow(a$stats),] )

# Add the labels
text(c(1:nlevels(mca$contex)) , DIM2_boxplot$stats[nrow(DIM2_boxplot$stats),]+over , LABELS[,1]  , col = my_colors[as.numeric(LABELS[,1])])

```

```{r MCA_platform_flakes, echo=FALSE}

# Filter specific categories for MCA analysis
platform_flake <- flake_tidy %>%
  select(Context, PlatformType, CortexPerc, ElongFact) %>%
  na.omit() %>%
  droplevels()

# Run MCA

mca_res_platform_flake <- MCA(platform_flake, graph = FALSE, quali.sup = 1, method = "Burt")

fviz_screeplot(mca_res_platform_flake, addlabels = TRUE, ylim = c(0, 45))

fviz_mca_var(mca_res_platform_flake, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE, invisible = "quali.sup",
             ggtheme = theme_gray())

```

```{r ANOVA_platform_elongated, echo=FALSE}

## From script tukey_boxplot

# Function below does not allow hyphen between categories

mca <- as.data.frame(mca_res_platform_flake$ind)
context <- as.data.frame(platform_flake$Context)

mca <- mca %>%
  select(1:2)

mca <- bind_cols(context, mca)

mca <- mca %>%
  mutate(context = `platform_flake$Context`, "dim1" = coord.Dim.1, "dim2" = coord.Dim.2) %>%
  select(context, dim1, dim2)

mca <- mca %>%
  mutate(context = dplyr::recode(context, "PAP 5'25-6'25" = "PAP 5'25 6'25",
                                 "PAP 4'75-5'25" = "PAP 4'75 5'25",
                                 "PAP 4'00-4'75" = "PAP 4'00 4'75"))

# DIMENSION 1

model <- lm(mca$dim1 ~ mca$context)
fit <- aov(model)
TUKEY <- TukeyHSD(x=fit, 'mca$context', conf.level=0.95)

generate_label_df <- function(TUKEY, variable){

# Extract labels and factor levels from Tukey post-hoc
Tukey.levels <- TUKEY[[variable]][,4]
Tukey.labels <- data.frame(multcompLetters(Tukey.levels)['Letters'])

# Put the labels in the same order as in the boxplot :
Tukey.labels$context=rownames(Tukey.labels)
Tukey.labels=Tukey.labels[order(Tukey.labels$context) , ]
return(Tukey.labels)
}

LABELS <- generate_label_df(TUKEY , "mca$context")

my_colors <- c( rgb(143,199,74,maxColorValue = 255),
             rgb(242,104,34,maxColorValue = 255),
             rgb(111,145,202,maxColorValue = 255),
             rgb(254,188,18,maxColorValue = 255),
             rgb(74,132,54,maxColorValue = 255),
             rgb(236,33,39,maxColorValue = 255),
             rgb(165,103,40,maxColorValue = 255))

# Draw the basic boxplot
DIM1_boxplot <- boxplot(mca$dim1 ~ mca$context , ylim=c(min(mca$dim1),
                                          1.1*max(mca$dim1)),
          col=my_colors[as.numeric(LABELS[,1])] , ylab="Dimension 1" , main="", outline = FALSE)

# I want to write the letter over each box. Over is how high I want to write it.
over <- 0.1*max(DIM1_boxplot$stats[nrow(DIM1_boxplot$stats),])

#Add the labels
text(c(1:nlevels(mca$context)), DIM1_boxplot$stats[nrow(DIM1_boxplot$stats),] +
       over, LABELS[,1] , col = my_colors[as.numeric(LABELS[,1])] )

# DIMENSION 2

model <- lm(mca$dim2 ~ mca$context)
fit <- aov(model)
TUKEY <- TukeyHSD(x=fit, 'mca$context', conf.level=0.95)

generate_label_df <- function(TUKEY, variable){

# Extract labels and factor levels from Tukey post-hoc
Tukey.levels <- TUKEY[[variable]][,4]
Tukey.labels <- data.frame(multcompLetters(Tukey.levels)['Letters'])

# Put the labels in the same order as in the boxplot
  Tukey.labels$context=rownames(Tukey.labels)
  Tukey.labels=Tukey.labels[order(Tukey.labels$context) , ]
  return(Tukey.labels)
}

LABELS <- generate_label_df(TUKEY , "mca$context")

my_colors <- c(rgb(143,199,74,maxColorValue = 255),
             rgb(242,104,34,maxColorValue = 255),
             rgb(111,145,202,maxColorValue = 255),
             rgb(254,188,18,maxColorValue = 255),
             rgb(74,132,54,maxColorValue = 255),
             rgb(236,33,39,maxColorValue = 255),
             rgb(165,103,40,maxColorValue = 255))

# Draw the basic boxplot
DIM2_boxplot <- boxplot(mca$dim2 ~ mca$context , ylim=c(min(mca$dim2),
                        1.1*max(mca$dim2)), col=my_colors[as.numeric(LABELS[,1])],                         ylab="Dimension 2", main="", outline = FALSE)

# Write the letter over each box. Over is how high I want to write it.
over <- 0.1*max( DIM2_boxplot$stats[nrow(a$stats),] )

# Add the labels
text(c(1:nlevels(mca$contex)) , DIM2_boxplot$stats[nrow(DIM2_boxplot$stats),]+over , LABELS[,1]  , col = my_colors[as.numeric(LABELS[,1])])

```