### 2019-09-18 analysis/paper/flakes_MCA.Rmd

---
title: "Elongated products"
author: 
  - name: João Cascalheira
    email: jmcascalheira@ualg.pt
    footnote: Corresponding author

address: 
  - code: ICArEHB
    address: | 
      FCHS, 
      Universidade do Algarve,
      Campus de Gambelas,
      8005-139 Faro, Portugal 
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: 
    rmarkdown::html_document:
        number_sections: yes
        toc: yes

---

```{r setup, include = FALSE, echo = FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>"
  )

library(knitr)
library(rmarkdown)
library(git2r)
library(readr)
library(dplyr)
library(tidyr)
library(tibble)
library(tab)
library(forcats)
library(FactoMineR)
library(factoextra)
library(multcompView)
library(RcmdrMisc)

```

```{r load_data, echo = FALSE}

# Load data, calculate Elongation/Flattening for blanks and tidy data (including lumping categories with little (5%) representation into 'Other')

blanks <- read.csv("../data/raw_data/blanks.csv")

flake_tidy <- blanks %>%
  filter(Class == "Flake") %>%
  drop_na(Length, Width, Thickness) %>%
  mutate(Elongation = Length / Width) %>%
  mutate(Flattening = Width / Thickness) %>%
  mutate(Profile = fct_lump(Profile, prop = 0.05)) %>%
  mutate(EdgeShape = fct_lump(EdgeShape, prop = 0.05)) %>%
  mutate(CrossSection = fct_lump(CrossSection, prop = 0.05)) %>%
  mutate(PlatformType = fct_lump(PlatformType, prop = 0.05)) %>%
  mutate(CortexLoc = fct_lump(CortexLoc, prop = 0.05)) %>%
  mutate(DorsalPattern = fct_lump(DorsalPattern, prop = 0.05)) %>%
  mutate(Termination = fct_lump(Termination, prop = 0.05)) %>%
  mutate(CortexPerc = dplyr::recode(CortexPerc, "75-95%" = "76-100%", "25-75%" ="26-75%", "100%" = "76-100%", "<25%" = "1-25%", ">95%" = "76-100%", "0%" = "0%")) %>%
  mutate(ScarNumber = as.factor(ScarNumber)) %>%
  mutate(DorsalPattern = dplyr::recode(DorsalPattern, "Unidentifiable" = "Other")) %>%
  mutate(ScarNumber = dplyr::recode(ScarNumber, "4" = "4 or more", "5" = "4 or more", "6" = "4 or more", "7" = "4 or more", "8" = "4 or more", "9" = "4 or more"))

# Rename Context column
colnames(flake_tidy)[1] <- "Context"

# Convert ScarNumber to factor
flake_tidy$ScarNumber <- factor(flake_tidy$ScarNumber)

# Bin continous variables
flake_tidy$ElongFact <- binVariable(flake_tidy$Elongation, bins = 3, method = "natural", labels = c("low", "medium", "high"))
flake_tidy$FlattFact <- binVariable(flake_tidy$Flattening, bins = 3, method = "natural", labels = c("low", "medium", "high"))

```

```{r flake_cross_table, echo=FALSE}

flake_table <- flake_tidy %>%
  droplevels()

# Set variables
var_list <- c("PlatformType", "CrossSection", "Profile", "DorsalPattern", "EdgeShape", "Elongation", "Flattening")

# Get cross-tables by chronology using cross_tb() function
flake_table <- tabmulti(flake_table, "Context", var_list,
                 p.include = FALSE,
                 n.headings = FALSE,
                 means.text.label = "none",
                 freq.text.label = "none")

flake_table <- flake_table %>% 
  as_tibble() %>% 
  select(" " = Variable, `AMB II`, `AMB IV`, `AMB VI`, `PAP 4'00-4'75`, `PAP 4'75-5'25`, `PAP 5'25-6'25`, `VALM`, `VB A`, `VB B`, `VB C`, Total = Overall)

kable(flake_table)

```

```{r MCA_core_exploitation_flake, echo=FALSE}

# Filter specific categories for MCA analysis
core_exploitation_flake <- flake_tidy %>%
  select(Context, DorsalPattern, CortexPerc, ScarNumber, ElongFact) %>%
  na.omit() %>%
  droplevels()

# Run MCA

mca_res_core_exploitation_flake <- MCA(core_exploitation_flake, graph = FALSE, quali.sup = 1, method = "Burt")

fviz_screeplot(mca_res_core_exploitation_flake, addlabels = TRUE, ylim = c(0, 45))

fviz_mca_var(mca_res_core_exploitation_flake, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE, invisible = "quali.sup",
             ggtheme = theme_gray())

```

# Variable contribution to MCA solution

```{r}

var_description_core_exploitation_flake <- data.frame(
  Dimension = c("Dimension 1", "Dimension 2"),
  Variability = c("34.5%", "17.9%"),
  Positive = c("Initial configuration of cores with different reduction strategies, high presence of cortex on dorsal surfaces, few removals, and low elongation of flakes",
    "Late moment of reduction sequence with no cortex on dorsal surfaces and high number of scars, using bidireccional or multidireccional strategies to extract highly elongated blanks"),
  Negative = c("Moment of the reduction sequence with small amounts of cortex, using bidireccional and unidireccional strategies for the extraction of highly elongated flakes",
  "Moment of the reduction sequence with high amounts of cortex, few removals, unidirectional strategies and low elongation of flakes"
))

kable(var_description_core_exploitation_flake)

```

```{r ANOVA_core_exploitation_elongated, echo=FALSE}

## From script tukey_boxplot

# Function below does not allow hyphen between categories

mca <- as.data.frame(mca_res_core_exploitation_flake$ind)
context <- as.data.frame(core_exploitation_flake$Context)

mca <- mca %>%
  select(1:2)

mca <- bind_cols(context, mca)

mca <- mca %>%
  mutate(context = `core_exploitation_flake$Context`, "dim1" = coord.Dim.1, "dim2" = coord.Dim.2) %>%
  select(context, dim1, dim2)

mca <- mca %>%
  mutate(context = dplyr::recode(context, "PAP 5'25-6'25" = "PAP 5'25 6'25",
                                 "PAP 4'75-5'25" = "PAP 4'75 5'25",
                                 "PAP 4'00-4'75" = "PAP 4'00 4'75"))

# DIMENSION 1

model <- lm(mca$dim1 ~ mca$context)
fit <- aov(model)
TUKEY <- TukeyHSD(x=fit, 'mca$context', conf.level=0.95)

generate_label_df <- function(TUKEY, variable){

# Extract labels and factor levels from Tukey post-hoc
Tukey.levels <- TUKEY[[variable]][,4]
Tukey.labels <- data.frame(multcompLetters(Tukey.levels)['Letters'])

# Put the labels in the same order as in the boxplot :
Tukey.labels$context=rownames(Tukey.labels)
Tukey.labels=Tukey.labels[order(Tukey.labels$context) , ]
return(Tukey.labels)
}

LABELS <- generate_label_df(TUKEY , "mca$context")

my_colors <- c( rgb(143,199,74,maxColorValue = 255),
             rgb(242,104,34,maxColorValue = 255),
             rgb(111,145,202,maxColorValue = 255),
             rgb(254,188,18,maxColorValue = 255),
             rgb(74,132,54,maxColorValue = 255),
             rgb(236,33,39,maxColorValue = 255),
             rgb(165,103,40,maxColorValue = 255))

# Draw the basic boxplot
DIM1_boxplot <- boxplot(mca$dim1 ~ mca$context , ylim=c(min(mca$dim1),
                                          1.1*max(mca$dim1)),
          col=my_colors[as.numeric(LABELS[,1])] , ylab="Dimesion 1" , main="", outline = FALSE)

# I want to write the letter over each box. Over is how high I want to write it.
over <- 0.1*max( a$stats[nrow(a$stats),] )

#Add the labels
text( c(1:nlevels(mca$contex)) , DIM1_boxplot$stats[nrow(DIM1_boxplot$stats),]+over , LABELS[,1]  , col=my_colors[as.numeric(LABELS[,1])] )

# DIMENSION 2

model <- lm(mca$dim2 ~ mca$context)
fit <- aov(model)
TUKEY <- TukeyHSD(x=fit, 'mca$context', conf.level=0.95)

generate_label_df <- function(TUKEY, variable){

# Extract labels and factor levels from Tukey post-hoc
Tukey.levels <- TUKEY[[variable]][,4]
Tukey.labels <- data.frame(multcompLetters(Tukey.levels)['Letters'])

# Put the labels in the same order as in the boxplot
  Tukey.labels$context=rownames(Tukey.labels)
  Tukey.labels=Tukey.labels[order(Tukey.labels$context) , ]
  return(Tukey.labels)
}

LABELS <- generate_label_df(TUKEY , "mca$context")

my_colors <- c(rgb(143,199,74,maxColorValue = 255),
             rgb(242,104,34,maxColorValue = 255),
             rgb(111,145,202,maxColorValue = 255),
             rgb(254,188,18,maxColorValue = 255),
             rgb(74,132,54,maxColorValue = 255),
             rgb(236,33,39,maxColorValue = 255),
             rgb(165,103,40,maxColorValue = 255))

# Draw the basic boxplot
DIM2_boxplot <- boxplot(mca$dim2 ~ mca$context , ylim=c(min(mca$dim2),
                                          1.1*max(mca$dim2)),
          col=my_colors[as.numeric(LABELS[,1])] , ylab="Dimension 2" , main="", outline = FALSE)

# Write the letter over each box. Over is how high I want to write it.
over <- 0.1*max(DIM2_boxplot$stats[nrow(a$stats),] )

# Add the labels
text(c(1:nlevels(mca$contex)) , DIM2_boxplot$stats[nrow(DIM2_boxplot$stats),]+over , LABELS[,1]  , col = my_colors[as.numeric(LABELS[,1])])

```

```{r MCA_platform_flakes, echo=FALSE}

# Filter specific categories for MCA analysis
platform_flake <- flake_tidy %>%
  select(Context, PlatformType, CortexPerc, ElongFact) %>%
  na.omit() %>%
  droplevels()

# Run MCA

mca_res_platform_flake <- MCA(platform_flake, graph = FALSE, quali.sup = 1, method = "Burt")

fviz_screeplot(mca_res_platform_flake, addlabels = TRUE, ylim = c(0, 45))

fviz_mca_var(mca_res_platform_flake, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE, invisible = "quali.sup",
             ggtheme = theme_gray())

```

# Variable contribution to MCA solution

```{r}

var_description_platform_flake <- data.frame(
  Dimension = c("Dimension 1", "Dimension 2"),
  Variability = c("18.4%", "12.7%"),
  Positive = c("Early moment in the reduction sequence, with high amounts of cortex on the core and percurssion platforms",
    "Early moment in the reduction sequence with high amounts of cortex on the core's debitage surface, with crushed and cortical platforms"),
  Negative = c("Late moment in sequence, with no cortex on dorsal surfaces, high elongation indices and crushed and faceted platforms",
  "Moment in sequence with significant amounts of cortical surface on the debitage surface, using faceted and plain platforms"
))

kable(var_description_platform_flake)

```

```{r ANOVA_platform_elongated, echo=FALSE}

## From script tukey_boxplot

# Function below does not allow hyphen between categories

mca <- as.data.frame(mca_res_platform_flake$ind)
context <- as.data.frame(platform_flake$Context)

mca <- mca %>%
  select(1:2)

mca <- bind_cols(context, mca)

mca <- mca %>%
  mutate(context = `platform_flake$Context`, "dim1" = coord.Dim.1, "dim2" = coord.Dim.2) %>%
  select(context, dim1, dim2)

mca <- mca %>%
  mutate(context = dplyr::recode(context, "PAP 5'25-6'25" = "PAP 5'25 6'25",
                                 "PAP 4'75-5'25" = "PAP 4'75 5'25",
                                 "PAP 4'00-4'75" = "PAP 4'00 4'75"))

# DIMENSION 1

model <- lm(mca$dim1 ~ mca$context)
fit <- aov(model)
TUKEY <- TukeyHSD(x=fit, 'mca$context', conf.level=0.95)

generate_label_df <- function(TUKEY, variable){

# Extract labels and factor levels from Tukey post-hoc
Tukey.levels <- TUKEY[[variable]][,4]
Tukey.labels <- data.frame(multcompLetters(Tukey.levels)['Letters'])

# Put the labels in the same order as in the boxplot :
Tukey.labels$context=rownames(Tukey.labels)
Tukey.labels=Tukey.labels[order(Tukey.labels$context) , ]
return(Tukey.labels)
}

LABELS <- generate_label_df(TUKEY , "mca$context")

my_colors <- c( rgb(143,199,74,maxColorValue = 255),
             rgb(242,104,34,maxColorValue = 255),
             rgb(111,145,202,maxColorValue = 255),
             rgb(254,188,18,maxColorValue = 255),
             rgb(74,132,54,maxColorValue = 255),
             rgb(236,33,39,maxColorValue = 255),
             rgb(165,103,40,maxColorValue = 255))

# Draw the basic boxplot
DIM1_boxplot <- boxplot(mca$dim1 ~ mca$context , ylim=c(min(mca$dim1),
                                          1.1*max(mca$dim1)),
          col=my_colors[as.numeric(LABELS[,1])] , ylab="Dimension 1" , main="", outline = FALSE)

# I want to write the letter over each box. Over is how high I want to write it.
over <- 0.1*max(DIM1_boxplot$stats[nrow(DIM1_boxplot$stats),])

#Add the labels
text(c(1:nlevels(mca$context)), DIM1_boxplot$stats[nrow(DIM1_boxplot$stats),] +
       over, LABELS[,1] , col = my_colors[as.numeric(LABELS[,1])] )

# DIMENSION 2

model <- lm(mca$dim2 ~ mca$context)
fit <- aov(model)
TUKEY <- TukeyHSD(x=fit, 'mca$context', conf.level=0.95)

generate_label_df <- function(TUKEY, variable){

# Extract labels and factor levels from Tukey post-hoc
Tukey.levels <- TUKEY[[variable]][,4]
Tukey.labels <- data.frame(multcompLetters(Tukey.levels)['Letters'])

# Put the labels in the same order as in the boxplot
  Tukey.labels$context=rownames(Tukey.labels)
  Tukey.labels=Tukey.labels[order(Tukey.labels$context) , ]
  return(Tukey.labels)
}

LABELS <- generate_label_df(TUKEY , "mca$context")

my_colors <- c(rgb(143,199,74,maxColorValue = 255),
             rgb(242,104,34,maxColorValue = 255),
             rgb(111,145,202,maxColorValue = 255),
             rgb(254,188,18,maxColorValue = 255),
             rgb(74,132,54,maxColorValue = 255),
             rgb(236,33,39,maxColorValue = 255),
             rgb(165,103,40,maxColorValue = 255))

# Draw the basic boxplot
DIM2_boxplot <- boxplot(mca$dim2 ~ mca$context , ylim=c(min(mca$dim2),
                        1.1*max(mca$dim2)), col=my_colors[as.numeric(LABELS[,1])],                         ylab="Dimension 2", main="", outline = FALSE)

# Write the letter over each box. Over is how high I want to write it.
over <- 0.1*max( DIM2_boxplot$stats[nrow(a$stats),] )

# Add the labels
text(c(1:nlevels(mca$contex)) , DIM2_boxplot$stats[nrow(DIM2_boxplot$stats),]+over , LABELS[,1]  , col = my_colors[as.numeric(LABELS[,1])])

```