### 2019-09-18 analysis/templates/template.Rmd

---
title: "My title"
author:
- João Cascalheira
- other author
date: '`r Sys.Date()`'
output: word_document
abstract: "My abstract \n"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Heading 1

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Heading 2

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

### Heading 3

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

#### Heading 4

Now knit this doc, and edit the styles in the resulting docx file. Then rename as 'template' and use this in the yml:

```
output:
  word_document:
    reference_docx: templates/template.docx
```

Don't forget line numbers for JAS! Headings are boldk, underline, nothing. And Arial Narrow 10pt for tables.

See here for more information: http://rmarkdown.rstudio.com/articles_docx.html