### 2019-11-18 analysis/paper/SOM3.Rmd

---
output: 
    bookdown::pdf_document2:
      toc: False
      number_sections: False
classoption: landscape

header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \pagenumbering{gobble}

---

```{r, setup, echo = FALSE, include=FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE)

library(tidyverse)
library(tab)
library(RcmdrMisc)
library(kableExtra)

```

```{r load_data, echo = FALSE}

# Load data
blanks <- read_csv("../data/raw_data/blanks.csv")
cores <- read_csv("../data/raw_data/cores.csv")

```

```{r cores, echo = FALSE}

# Calculate Elongation/Flattening for blanks and tidy data (including lumping categories with little (5%) representation into 'Other')
cores_tidy <- cores %>%
  drop_na(Length, Width, Thickness) %>% 
  mutate(Elongation = Length / Width) %>%
  mutate(Flattening = Width / Thickness) %>% 
  mutate(CoreType = fct_lump(CoreType, prop = 0.05)) %>%
  mutate(BlankType = fct_lump(BlankType, prop = 0.05)) %>%
  mutate(CoreSection = fct_lump(CoreSection, prop = 0.05)) %>%
  mutate(PlatformType = fct_lump(PlatformType, prop = 0.05)) %>%
  mutate(CortexLoc = fct_lump(CortexLoc, prop = 0.05)) %>%
  mutate(CortexPerc = dplyr::recode(CortexPerc, "75-95%" = "76-100%", "25-75%" ="26-75%", "100%" = "76-100%", "<25%" = "1-25%", ">75%" = "76-100%", "0%" = "0%")) %>% 
  mutate(BlankType = dplyr::recode(BlankType, "Blades" = "Elongated product", "Bladelets" = "Elongated product"))

# Bin continous variables ("natural" is used to cut between bins to be determined by a k-means clustering)
cores_tidy$ElongFact <- binVariable(cores_tidy$Elongation, bins = 2, method = "natural", labels = c("low", "high"))
cores_tidy$FlattFact <- binVariable(cores_tidy$Flattening, bins = 2, method = "natural", labels = c("low", "high"))
cores_tidy$WeightFact <- binVariable(cores_tidy$Weight, bins = 2, method = "natural", labels = c("low", "high"))

```

## Table S3 - Cores attribute frequency

```{r cores_cross_table, echo=FALSE}

var_list <- c("CoreType", "Platforms", "CoreSection", "BlankType", "CortexPerc", "Weight", "Elongation", "Flattening")

cores_tidy <- as.data.frame(cores_tidy)

cores_tidy$Platforms <- as.factor(cores_tidy$Platforms)

cores_table <- tabmulti(cores_tidy, "Context", var_list,
                 p.include = FALSE,
                 n.headings = FALSE,
                 bold.varnames = TRUE,
                 bold.colnames = TRUE,
                 ynames = c(CoreType = "Core type", Platforms = "Platforms", CoreSection = "Core section", BlankType = "Blank type", CortexPerc = "Cortex %", Weight = "Weight", Elongation = "Elongation", Flattening = "Flattening"))

cores_table <- cores_table %>% 
  as_tibble() %>% 
  select(" " = Variable, `AMB II`, `AMB IV`, `AMB VI`, `PAP 4'00-4'75`, `PAP 4'75-5'25`, `PAP 5'25-6'25`, `VALM`, `VB A`, `VB B`, `VB C`, Total = Overall)

knitr::kable(cores_table, "latex", booktabs = TRUE) %>% 
  kable_styling(latex_options = c("scale_down"))

```